>

### 2019-04-22 paper/spinifex-paper.tex
first two sections of algorithm section

The section below describes the algorithm for performing a 2D radial
manual tour:

@@

These algorithm steps are described in more detail below with
descriptions, equations, and figures are given inline. Function names
are mentioned, but code usage is discussed in more detail in section
\ref{sec:usage}.

@@

\begin{itemize}
  \item $\textbf{X}$, the data, an $n \times p$ numeric matrix to be embedded in two dimensions
  \item $\textbf{B} = (B_1,~ B_2)$, any orthonormal projection basis set, $p \times 2$ matrix, describing the projection from $p$ to two dimensions
  \item $\textbf{e}$, a zero column vector of length $p$ with the $k-$th element set to one, where $k$ is the number of the variable to manipulate, the manip var
  \item $\theta_i$, the sequence of angles of in-projection-plane rotation, for example, on the reference axes
  \item $\phi_i$, the sequence of angles of out-of-projection-plane rotation, coming into the manipulation space
\end{itemize}

@@

\hypertarget{toy-data-set}{%
\subsection{Toy data set}\label{toy-data-set}}

@@

The data is defined. A basis set, the orientation of the projection to
explore, must be provided. One way to identify a projection containing
interesting features is to apply a guided tour. In a guided tour, the
projection sequence is selected by optimizing an index function via
hill-climbing on the projection space. In this case, the holes index is
selected and applied to standardized flea data. The holes index is
maximized when the projected observations are furthest from the center.
Figure \ref{fig:step0} shows a locally optimized projection of the data.
The left panel displays the reference axes of the projection basis, a
visual indication of the magnitude and direction each variable
contributes to the projection. The right panel shows the data as
projected through the basis set described by the reference axes (left).
Data points are colored and given shape according to the species (while
the guided tour was unsupervised with this information).

@@

\caption[Basis reference axes (left) and projected data (right) of standardized flea data]{Basis reference axes (left) and projected data (right) of standardized flea data. Data points color and shape are mapped to beetle species. The basis was identified by a holes-index guided tour. The variables aede2 and tars1 contribute mostly orthogonal to the other variables.}\label{fig:step0}
\end{figure}
\end{Schunk}

Call \texttt{view\_basis()} on a basis to produce a \CRANpkg{ggplot2}
(Wickham \protect\hyperlink{ref-wickham_ggplot2:_2016}{2016}) graphic
similar to figure \ref{fig:step0}. Projection space is always available
for display via the matrix multiplication
\(\textbf{X}_{[n,~p]} ~*~ \textbf{B}_{[p,~d]} ~=~ \textbf{P}_{[n,~d]}\).

>>>>

This section describes the algorithm for performing a 2D radial manual
tour, containing these steps:

@@

More detail is provided in the following sections. Function names are
provided, but code structure and usage is discussed in section
\ref{sec:usage}.

@@

\begin{itemize}
\tightlist
\item
  \(\textbf{X}\), the data, an \(n \times p\) numeric matrix to be
  projected to 2D.
\item
  \(\textbf{B} = (B_1,~ B_2)\), any 2D orthonormal projection basis,
  \(p \times 2\) matrix, describing the projection from \(p\) to two
  dimensions
\item
  \(\textbf{e}\), a 1D coordinate basis vector of length \(p\) with the
  \(k-\)th element set to one, where \(k\) is the index of the variable
  to manipulate, called the ``manip var''.
\item
  \(\theta_i\), the sequence of angles of in-projection-plane rotation,
  for example, on the reference axes.
\item
  \(\phi_i\), the sequence of angles of out-of-projection-plane
  rotation, coming into the manipulation space.
\end{itemize}

@@

\hypertarget{set-up}{%
\subsection{Set up}\label{set-up}}

@@

An initial 2D projection basis set must be provided. One way to identify
a projection containing interesting features is to apply a guided tour.
In a guided tour, the projection sequence is selected by optimizing an
index function via hill-climbing on the projection space. In this case,
the holes index is selected and applied to standardized flea data. The
holes index is maximized when the projected observations are furthest
from the center. Figure \ref{fig:step0} shows a locally optimized
projection of the data. The left panel displays the reference axes of
the projection basis, a visual indication of the magnitude and direction
each variable contributes to the projection. The right panel shows the
data as projected through the basis set described by the reference axes
(left). Data points are colored and given shape according to the
species.

Call \texttt{view\_basis()} on a basis to produce a \CRANpkg{ggplot2}
(Wickham \protect\hyperlink{ref-wickham_ggplot2:_2016}{2016}) graphic
similar to Figure \ref{fig:step0}. The projected data is always
available for display via matrix multiplication
\(\textbf{X}_{[n,~p]} ~*~ \textbf{B}_{[p,~d]} ~=~ \textbf{P}_{[n,~d]}\).

@@

\caption[Initial 2D projection]{Initial 2D projection: representation of the basis  (left) and resulting data projection (right) of standardized flea data. The color and shape of data points are mapped to beetle species. The basis was identified using a projection pursuit guided tour, with the holes index. The contribution of the variables aede2 and tars1 approximately contrasts the other variables. The visible structure in the projection are the three clusters corresponding to the three species.}\label{fig:step0}
\end{figure}
\end{Schunk}

### 2019-04-22 paper/spinifex_paper.Rmd
first two sections of algorithm section

<!-- remember mixing latex and rmarkdown bullets isn't trivial. -->
\begin{itemize}
  \item $\textbf{X}$, the data, an $n \times p$ numeric matrix to be embedded in two dimensions
  \item $\textbf{B} = (B_1,~ B_2)$, any orthonormal projection basis set, $p \times 2$ matrix, describing the projection from $p$ to two dimensions
  \item $\textbf{e}$, a zero column vector of length $p$ with the $k-$th element set to one, where $k$ is the number of the variable to manipulate, the manip var
  \item $\theta_i$, the sequence of angles of in-projection-plane rotation, for example, on the reference axes
  \item $\phi_i$, the sequence of angles of out-of-projection-plane rotation, coming into the manipulation space
\end{itemize}

@@

## Toy data set

@@

<!-- Projection basis -->
The data is defined. A basis set, the orientation of the projection to explore, must be provided. One way to identify a projection containing interesting features is to apply a guided tour. In a guided tour, the projection sequence is selected by optimizing an index function via hill-climbing on the projection space. In this case, the holes index is selected and applied to standardized flea data. The holes index is maximized when the projected observations are furthest from the center. Figure \ref{fig:step0} shows a locally optimized projection of the data. The left panel displays the reference axes of the projection basis, a visual indication of the magnitude and direction each variable contributes to the projection. The right panel shows the data as projected through the basis set described by the reference axes (left). Data points are colored and given shape according to the species (while the guided tour was unsupervised with this information).

```{r step0, echo=F, results='hide', out.width='70%',                             fig.cap = "Basis reference axes (left) and projected data (right) of standardized flea data. Data points color and shape are mapped to beetle species. The basis was identified by a holes-index guided tour. The variables aede2 and tars1 contribute mostly orthogonal to the other variables."}
## Flea holes tour
f_dat  <- tourr::rescale(flea[,1:6])
f_cat  <- factor(flea$species)
f_path <- save_history(f_dat, guided_tour(holes()))
f_bas  <- matrix(f_path[,, max(dim(f_path)[3])], ncol=2)
f_mvar <- 5
f_msp  <- create_manip_space(basis = f_bas, manip_var = f_mvar)
f_proj <- data.frame(tourr::rescale(f_dat %*% f_msp[, 1:2]))

@@

Call `view_basis()` on a basis to produce a \CRANpkg{ggplot2} [@wickham_ggplot2:_2016] graphic similar to figure \ref{fig:step0}. Projection space is always available for display via the matrix multiplication $\textbf{X}_{[n,~p]} ~*~ \textbf{B}_{[p,~d]} ~=~ \textbf{P}_{[n,~d]}$.

>>>>

<!-- remember mixing latex and rmarkdown bullets isn't trivial. -->
- $\textbf{X}$, the data, an $n \times p$ numeric matrix to be projected to 2D.
- $\textbf{B} = (B_1,~ B_2)$, any 2D orthonormal projection basis, $p \times 2$ matrix, describing the projection from $p$ to two dimensions
- $\textbf{e}$, a 1D coordinate basis vector of length $p$ with the $k-$th element set to one, where $k$ is the index of the variable to manipulate, called the "manip var".
- $\theta_i$, the sequence of angles of in-projection-plane rotation, for example, on the reference axes.
- $\phi_i$, the sequence of angles of out-of-projection-plane rotation, coming into the manipulation space.

@@

## Set up

@@

<!-- Projection basis -->
An initial 2D projection basis set must be provided. One way to identify a projection containing interesting features is to apply a guided tour. In a guided tour, the projection sequence is selected by optimizing an index function via hill-climbing on the projection space. In this case, the holes index is selected and applied to standardized flea data. The holes index is maximized when the projected observations are furthest from the center. Figure \ref{fig:step0} shows a locally optimized projection of the data. The left panel displays the reference axes of the projection basis, a visual indication of the magnitude and direction each variable contributes to the projection. The right panel shows the data as projected through the basis set described by the reference axes (left). Data points are colored and given shape according to the species.

Call `view_basis()` on a basis to produce a \CRANpkg{ggplot2} [@wickham_ggplot2:_2016] graphic similar to Figure \ref{fig:step0}. The projected data is always available for display via matrix multiplication $\textbf{X}_{[n,~p]} ~*~ \textbf{B}_{[p,~d]} ~=~ \textbf{P}_{[n,~d]}$.

```{r step0, echo=F, results='hide', out.width='70%',                             fig.cap = "Initial 2D projection: representation of the basis  (left) and resulting data projection (right) of standardized flea data. The color and shape of data points are mapped to beetle species. The basis was identified using a projection pursuit guided tour, with the holes index. The contribution of the variables aede2 and tars1 approximately contrasts the other variables. The visible structure in the projection are the three clusters corresponding to the three species."}
## Flea holes tour
f_dat  <- tourr::rescale(flea[,1:6])
f_cat  <- factor(flea$species)
f_path <- save_history(f_dat, guided_tour(holes()))
f_bas  <- matrix(f_path[,, max(dim(f_path)[3])], ncol=2)
f_mvar <- 5
f_msp  <- create_manip_space(basis = f_bas, manip_var = f_mvar)
f_proj <- data.frame(tourr::rescale(f_dat %*% f_msp[, 1:2]))

@@
